### 2016-07-11 Summary.tex
some changes

To this end a model file, generated by \texttt{Sarah}, which serves only for leading order processes has been adjusted manually to incorporate counter term Feynman rules. At this point, Philip Dießner did a major contribution when the model file was changed. After this, the model file has been provided with appropriate renormalization constants which are explicitly given in the thesis. This includes also a supersymmetry restoring counterterm. By means of this model file, renormalized one-loop corrections to the matrix element of squark production has been calculated in \texttt{Mathematica}, using the packages \texttt{FeynArts} and \texttt{FormCalc}. After this, the virtual contributions, which have been checked for ultraviolet-finiteness have been passed to a \texttt{C++} code and have been evaluated using the \texttt{Looptools} package. Finally, the phase space integration has been performed numerically using the \texttt{CUBA} library for parallelized numerical integration. However, the virtual corrections to the cross section are not infrared finite. To render the next-to-leading order cross section $\sigma^{\mathrm{NLO}}$ finite, real corrections need to be included. This has been done by Wojciech Kotlarski. After including them, $\sigma^{\mathrm{NLO}}$ has been found to contain no further divergences.\\
Equipped with this programm, the $K$-factors for up-squark production have been calculated as a function of the squark mass and a function of the gluino mass. As a result, the comparison of $K$-factors for squark production in the MRSSM and the MSSM showed that the very same are slightly larger in the MRSSM. But since the tree-level cross section of squark production in the MRSSM are, depending on the choice  of squark and gluino mass, about an order of magnitude smaller than in the MSSM, this renders the MRSSM to a perfectly viable extension of the Standard Model.\\

>>>>

To this end a model file, generated by \texttt{Sarah}, which serves only for leading order processes has been adjusted manually to incorporate counter term Feynman rules. At this point, Philip Dießner did a major contribution when the model file was changed. After this, the model file has been provided with appropriate renormalization constants which are explicitly given in the thesis. This includes also a supersymmetry restoring counterterm. By means of this model file, renormalized one-loop corrections to the matrix element of squark production has been calculated in \texttt{Mathematica}, using the packages \texttt{FeynArts} and \texttt{FormCalc}. After this, the virtual contributions, which have been checked for ultraviolet-finiteness have been passed to a \texttt{C++} code and have been evaluated using the \texttt{Looptools} package. Finally, the phase space integration has been performed numerically using the \texttt{CUBA} library for parallelized numerical integration. However, the virtual corrections to the cross section are not infrared finite. To render the next-to-leading order cross section $\sigma^{\mathrm{NLO}}$ finite, real corrections need to be included. This has been done by Wojciech Kotlarski. After including them, $\sigma^{\mathrm{NLO}}$ has been found to contain no further divergences.
\begin{table}[H]
\begin{center}
\begin{tabular}{c|c?c|c|c?c|c|c}
\multicolumn{2}{c?}{} & \multicolumn{3}{c?}{MRSSM: $p + p \to \tilde{u}_L + \tilde{u}_R$} & \multicolumn{3}{c}{MSSM: $p + p \to \tilde{u} + \tilde{u}$}\\
\hlinewd{2pt}
$m_{\tilde{q}}$ in GeV & $m_{\tilde{g}}$ in GeV & $\sigma^{\mathrm{LO}}$ in fb & $\sigma^{\mathrm{NLO}}$ in fb & $K$ & $\sigma^{\mathrm{LO}}$ in fb & $\sigma^{\mathrm{NLO}}$ in fb & $K$\\
\hlinewd{2pt}
$500$ & $500$ & $966.4$ & $1441$ & $1.49$ & $2333$ & $3102$ & $1.33$\\
$500$ & $1000$ & $303.4$ & $383$ & $1.26$ & $1271$ & $1393$ & $1.10$\\
$1000$ & $1000$ & $42.63$ & $61.8$ & $1.45$ & $122.2$ & $161.3$ & $1.32$\\
$1000$ & $2000$ & $11.56$ & $14.3$ & $1.24$ & $6.36$ & $72.1$ & $1.09$\\
$1500$ & $3000$ & $0.9166$ & $1.13$ & $1.23$ & $7.045$ & $7.57$ & $1.07$
\end{tabular}
\caption{Total hadronic cross section for the production of up-squarks in the MRSSM and the MSSM at leading order and next-to-leading order. Also given are the $K$-factors. As a consequence of $R$-charge conservation, the only allowed channel in the MRSSM is $\tilde{u}_L + \tilde{u}_R$ production, whereas in the MSSM also $\tilde{u}_L + \tilde{u}_L$ and $\tilde{u}_R + \tilde{u}_R$ production is allowed.\newline 
The results are shown for a selected set of masses. The center-of-mass energy is $\sqrt{S} = \unit[13]{TeV}$ and within the MRSSM the pseudoscalar mass is fixed to $m_{\mathrm{\sigma}} = \unit[5000]{GeV}$.}\label{tab:MRSSMKfactors}
\end{center}
\end{table}
Equipped with this program, the $K$-factors for up-squark production have been calculated as a function of the squark mass and a function of the gluino mass. As a result, the comparison of $K$-factors for squark production in the MRSSM and the MSSM showed that the very same are slightly larger in the MRSSM. But since the tree-level cross section of squark production in the MRSSM are, depending on the choice  of squark and gluino mass, about an order of magnitude smaller than in the MSSM, this renders the MRSSM to a perfectly viable extension of the Standard Model.\\

### 2016-07-11 Virt_Real.tex
some changes

\section{Virtual and Real Corrections}\label{sec:VirtRealCorr}
The results calculated in the previous chapter are not the full theoretical predictions for the cross sections of the different processes but only the first term in a perturbation series of $\alpha_s$.\\
This section describes the necessary steps in the calculation of the squark production cross section at next-to-leading order, i.e. its $\mathcal{O}(\alpha_s)$-correction.\\
\ignore{Reason validity of perturbation theory in this context.}

@@

& \phi^0 / \sigma^0 \to \sqrt{Z_{\phi^0 / \sigma^0}}\phi^0 / \sigma^0 && \tilde{q}_{L/R} \to \sqrt{Z_{\tilde{q}_{L/R}}}\tilde{q}_{L/R}\nonumber\\
& P_L \tilde{g} \to \sqrt{Z_{\tilde{g}}^L} P_L \tilde{g} && P_R \tilde{g} \to \sqrt{Z_{\tilde{g}}^R} P_R \tilde{g}\nonumber\\
& q \to \sqrt{Z_q}q && G_\mu \to \sqrt{Z_G} G_\mu && \label{eq:fieldtrafo}\\
& g_s \to g_{s\ \mathrm{bare}} && m_{\tilde{q}}^2 \to m_{\tilde{q}\ \mathrm{bare}}^2\nonumber\\
&  m_{\sigma}^2 \to m_{\sigma\ \mathrm{bare}}^2 && m_{\tilde{g}} \to m_{\tilde{g}\ \mathrm{bare}}\label{eq:parametertrafo}

@@

where $\sigma^{\mathrm{R}}$ donates the cross section arising from real corrections. However, real corrections also contain divergences. To see how these arise, consider the second diagram in fig. \ref{fig:RealGluon}. The propagator of the quark which radiates off a gluon gives a contribution of the form

@@

\subsubsection{Real Quark Radiation}
As already mentioned before, adding up virtual and real gluon emission contributions results in the cancellation of divergences. However, there is a further contribution to squark production at next-to-leading order. This involves the radiation of an additional antiquark in the final state. To meet color conservation, the initial state has to be composed of a quark and a gluon, see fig. \ref{fig:quarkradiation}.
\begin{figure}[!htbp]
\begin{center}
\begin{tikzpicture}[line width=1.5 pt, scale=1.3]
\begin{scope}[shift={(0,0.25)}]
	\draw[fermion] (-1.25,0.75)--(-0.5,0);
	\draw[gluon] (-1.25,-0.75)--(-0.5,0);
	\draw[fermion] (-0.5,0)--(0.5,0); 
	\draw[scalar] (0.5,0)--(1.25,0.75);
	\draw[gluon] (0.5,0)--(1.25,-0.75);
	\draw[fermionnoarrow] (0.5,0)--(1.25,-0.75);
	\draw[scalar] (1.25,-0.75)--(2,0);
	\draw[fermion] (2,-1.5)--(1.25,-0.75);
	\end{scope}
\begin{scope}[shift={(3,0)}]
	\draw[fermion] (0,1)--(1,1);	
	\draw[gluon] (0,-0.5)--(1,-0.5);
	\draw[fermionnoarrow] (1,1)--(1,-0.5);
	\draw[gluon] (1,1)--(1,-0.5);
	\draw[scalar] (1,1)--(2.75,1);
	\draw[fermionnoarrow] (1,-0.5)--(2,-0.5);
	\draw[gluon] (1,-0.5)--(2,-0.5);
	\draw[scalar] (2,-0.5)--(2.75,0.25);
	\draw[fermion] (2.75,-1.25)--(2,-0.5);
\end{scope}
\begin{scope}[shift={(7,0)}]
	\draw[fermion] (0,1)--(1,1);
	\draw[gluon] (0,-1)--(1,-1);
	\draw[fermionnoarrow] (1,1)--(1,0);
	\draw[gluon] (1,1)--(1,0);
	\draw[fermion] (1,-1)--(1,0);
	\draw[scalar] (1,1)--(2,1);
	\draw[fermion] (2,-1)--(1,-1);
	\draw[scalar] (1,0)--(2,0);
\end{scope}
\end{tikzpicture}
\caption{A selection of real quark emission diagrams contributing to squark production. The first two diagrams have a gluino propagator which may go on-shell. In this regime of phase space these diagrams should not be included in the correction to squark-production but to a tree-level contribution to squark-gluino-production.}\label{fig:quarkradiation}
\end{center}
\end{figure}
These contributions do not comprise soft divergences. A split-up of space space in a soft and a hard part is therefore not needed. However, there are collinear singularities, originating from initial state radiation like shown in the last diagram of \ref{fig:quarkradiation}, which again have to be removed by means of mass factorization.\\
But there is another complication: The first and second diagram in fig. \ref{fig:quarkradiation} include a gluino which subsequently decays into a squark-antiquark pair. If the squared gluino's four-momentum approaches its mass-shell $p^2 = m^2_{\tilde{g}}$ the amplitude of this diagram hits a singularity. Of course, this can only happen if $m_{\tilde{g}} > m_{\tilde{q}}$.\\
These singularities are treated by introducing a non-zero width $\Gamma_{\tilde{q}}$ for the gluino as a regulator, i.e. modifying the propagator to the Breit-Wigner form
\begin{align}
\frac{1}{p^2 - m_{\tilde{q}}^2} \to \frac{1}{p^2 - m_{\tilde{q}}^2 + i m_{\tilde{q}} \Gamma_{\tilde{q}}}.
\end{align}
Integrating over the regime of phase space, where the gluino approaches its mass-shell, actually gives the tree-level contribution of the process $qG \to \tilde{q}\tilde{g}$ weighted by the branching ratio $\frac{\Gamma_{\tilde{q}\to q \tilde{g}}}{\Gamma_{\tilde{q}}}$. To avoid double counting, one has to cut this very regime of phase space out and subtract it from the total contribution originating from those diagrams. For a detailed prescription on how to deal with this issue, see \cite{Beenakker:1996ch, Gavin:2013kga}.

>>>>

\section{Virtual and Real Corrections}\label{sec:VirtRealCorr}
The results calculated in the previous chapter are not the full theoretical predictions for the cross sections of the various processes but only the first term in a perturbation series of $\alpha_s$.\\
To provide accurate predictions, it is necessary that correction to the leading order are small. This implies that $\alpha_s$, which depends strongly on the energy scale, needs to fulfill $\alpha_s < 1$. It is known that $\alpha_s$ becomes of $\mathcal{O}(1)$ and larger when the energy scale is of $\mathcal{O}(\unit[200]{MeV})$ and smaller.\cite{dissertori2003quantum} Fortunately squark production is a process which involves large enough energy to maintain the validity of perturbation theory as the typical energy scale involved, is the mass of squarks which certainly fulfills $m_{\tilde{q}} > \unit[500]{GeV}$.\\
This section describes the necessary steps in the calculation of the squark production cross section at next-to-leading order, i.e. its $\mathcal{O}(\alpha_s)$-correction.\\
\ignore{Reason validity of perturbation theory in this context.}

@@

& \phi^0 / \sigma^0 \to \sqrt{Z_{\phi^0 / \sigma^0}}\phi^0 / \sigma^0, && \tilde{q}_{L/R} \to \sqrt{Z_{\tilde{q}_{L/R}}}\tilde{q}_{L/R}, \nonumber\\
& P_L \tilde{g} \to \sqrt{Z_{\tilde{g}}^L} P_L \tilde{g}, && P_R \tilde{g} \to \sqrt{Z_{\tilde{g}}^R} P_R \tilde{g},\nonumber\\
& q \to \sqrt{Z_q}q && G_\mu \to \sqrt{Z_G} G_\mu, && \label{eq:fieldtrafo}\\
& g_s \to g_{s\ \mathrm{bare}}, && m_{\tilde{q}}^2 \to m_{\tilde{q}\ \mathrm{bare}}^2, \nonumber\\
&  m_{\sigma}^2 \to m_{\sigma\ \mathrm{bare}}^2, && m_{\tilde{g}} \to m_{\tilde{g}\ \mathrm{bare}}.\label{eq:parametertrafo}

@@

where $\sigma^{\mathrm{R}}$ denotes the cross section arising from real corrections. However, real corrections also contain divergences. To see how these arise, consider the second diagram in fig. \ref{fig:RealGluon}. The propagator of the quark which radiates off a gluon gives a contribution of the form

@@

\subsubsection{Real Quark Radiation}\label{sec:RealQuarkRad}
As already mentioned before, adding up virtual and real gluon emission contributions results in the cancellation of divergences. However, there is a further contribution to squark production at next-to-leading order. This involves the radiation of an additional antiquark in the final state. To meet color conservation, the initial state has to be composed of a quark and a gluon, see fig. \ref{fig:quarkradiation}.
\begin{figure}[!htbp]
\begin{center}
\begin{tikzpicture}[line width=1.5 pt, scale=1.3]
\begin{scope}[shift={(0,0.25)}]
	\draw[fermion] (-1.25,0.75)--(-0.5,0);
	\draw[gluon] (-1.25,-0.75)--(-0.5,0);
	\draw[fermion] (-0.5,0)--(0.5,0); 
	\draw[scalar] (0.5,0)--(1.25,0.75);
	\draw[gluon] (0.5,0)--(1.25,-0.75);
	\draw[fermionnoarrow] (0.5,0)--(1.25,-0.75);
	\draw[scalar] (1.25,-0.75)--(2,0);
	\draw[fermion] (2,-1.5)--(1.25,-0.75);
	\end{scope}
\begin{scope}[shift={(3,0)}]
	\draw[fermion] (0,1)--(1,1);	
	\draw[gluon] (0,-0.5)--(1,-0.5);
	\draw[fermionnoarrow] (1,1)--(1,-0.5);
	\draw[gluon] (1,1)--(1,-0.5);
	\draw[scalar] (1,1)--(2.75,1);
	\draw[fermionnoarrow] (1,-0.5)--(2,-0.5);
	\draw[gluon] (1,-0.5)--(2,-0.5);
	\draw[scalar] (2,-0.5)--(2.75,0.25);
	\draw[fermion] (2.75,-1.25)--(2,-0.5);
\end{scope}
\begin{scope}[shift={(7,0)}]
	\draw[fermion] (0,1)--(1,1);
	\draw[gluon] (0,-1)--(1,-1);
	\draw[fermionnoarrow] (1,1)--(1,0);
	\draw[gluon] (1,1)--(1,0);
	\draw[fermion] (1,-1)--(1,0);
	\draw[scalar] (1,1)--(2,1);
	\draw[fermion] (2,-1)--(1,-1);
	\draw[scalar] (1,0)--(2,0);
\end{scope}
\end{tikzpicture}
\caption{A selection of real quark emission diagrams contributing to squark production. The first two diagrams have a gluino propagator which may go on-shell. In this regime of phase space these diagrams should not be included in the correction to squark-production but to a tree-level contribution to squark-gluino-production.}\label{fig:quarkradiation}
\end{center}
\end{figure}
These contributions do not comprise soft divergences, as they are associated with a different initial state and the factorization theorem guarantees the absences of soft divergences in each initial state separately. In other words, real correction, coming from an initial state with quark and gluon, cannot contribute to the cancellation of soft divergences with virtual correction which are associated with two quarks in the initial state.\\
A split-up of space space in a soft and a hard part is therefore not needed. However, there are collinear singularities, originating from initial state radiation like shown in the last diagram of \ref{fig:quarkradiation}, which again have to be removed by means of mass factorization.\\
But there is another complication: The first and second diagram in fig. \ref{fig:quarkradiation} include a gluino which subsequently decays into a squark-antiquark pair. If the squared gluino's four-momentum approaches its mass-shell $p^2 = m^2_{\tilde{g}}$ the amplitude of this diagram hits a singularity. Of course, this can only happen if $m_{\tilde{g}} > m_{\tilde{q}}$.\\
These singularities are treated by introducing a non-zero width $\Gamma_{\tilde{q}}$ for the gluino as a regulator, i.e. modifying the propagator to the Breit-Wigner form
\begin{align}
\frac{1}{p^2 - m_{\tilde{q}}^2} \to \frac{1}{p^2 - m_{\tilde{q}}^2 + i m_{\tilde{q}} \Gamma_{\tilde{q}}}.
\end{align}
Integrating over the regime of phase space, where the gluino approaches its mass-shell, actually gives the tree-level contribution of the process $qG \to \tilde{q}\tilde{g}$ weighted by the branching ratio $\frac{\Gamma_{\tilde{q}\to q \tilde{g}}}{\Gamma_{\tilde{q}}}$. To avoid double counting, one has to cut this very regime of phase space out and subtract it from the total contribution originating from those diagrams. For a detailed prescription on how to deal with this issue, see \cite{Beenakker:1996ch, Gavin:2013kga}.
